---
title: "Exam 3"
author: "Tom Wilson"
date: "Dember 12th, 2018"
output: pdf_document
latex_engine: xelatex
---

 *Required R packages*
```{r warning=FALSE, results='hide', message=FALSE} 
library(tidyverse)
library(data.table)
library(glmnet)
library(glmnetUtils)
library(MASS)
library(caret)
library(leaps)
library(boot)
library(knitr)
```

# 1.
Predict the number of applications received using the other variables in the college data set. 
```{r}
college <- fread('../data/College.csv') %>% subset(,-V1)
```

## a.
Split the data set into a training set and a test set.


```{r}
n=nrow(college)
train_sample <- runif(n,0,1) > 1 - 0.75 #random uniform sample
college_train <- college[ train_sample,] %>% select_if(is.numeric)
college_test  <- college[!train_sample,] %>% select_if(is.numeric)
```


```{r}
x_train <- college_train %>% subset(,-Apps) # all columns except Apps
y_train <- college_train %>% subset(, Apps) 
x_test  <- college_test  %>% subset(,-Apps) 
y_test  <- college_test  %>% subset(, Apps)
```

## b.
Fit a linear model using least squares on the training set, and report the test error obtained.


```{r}
linear_model <- lm(formula = Apps~., data = college_train )
```


```{r}
residual <- predict(linear_model,newdata = college_test) - college_test$Apps
RMSE <- sqrt(sum(residual^2))
```

Using linear regression, Root Mean Squared Error on test dataset is

```{r echo=FALSE}
RMSE %>% round(2)
```

### c.
Fit a ridge regression model on the training set, with $\lambda$ chosen by cross-validation.  Report the test error obtained.

using glmnetUtils
```{r}
fit <- cv.glmnet(Apps ~ ., data=college_train)
plot(fit)
fit
```

MSE is minimized at $\lambda$ of

```{r echo=FALSE}
fit$lambda.min
```


```{r}
best.fit <- glmnet(Apps ~ ., data=college_train, lambda=fit$lambda.min)
pred <- predict(best.fit,newdata=college_test)
residual <- pred - college_test$Apps
RMSE <- sqrt(sum(residual^2))
```

Using ridge regression, Root Mean Squared Error on test dataset is

```{r echo=FALSE}
RMSE %>% round(2)
```

Which is slightly worse than ordinary least squares shown previously.

## 2. 
Consider the Boston housing data set, from the MASS library.

```{r}
Boston %>% head() %>% kable()
```

### a.
Based on this data set, provide an estimate for the population mean of medv.  Call this estimate.

```{r}
mu_hat <- mean(Boston$medv)
```

An estimate for the population mean of medv, $\hat{\mu}$ is
```{r}
round(mu_hat,4)
```

### b.
Provide an estimate of the standard error of   Interpret this result.
Hint:  We can compute the standard error of the sample mean by dividing the sample standard deviation by the square root of the number of observations.


```{r}
n <- nrow(Boston)
SE <- sd(Boston$medv)/sqrt(n)
```

an estimate of the standard error of the population mean of medv is

```{r echo=FALSE}
round(SE,4)
```

### c.
Now estimate the standard error of  using the bootstrap.  How does this compare to your answer from (b)?

```{r}
SE.fn=function(data,index){
    n=length(data[index])
 return(sd(data[index])/sqrt(n))
 }

SE_bootstrap <- boot(data = Boston$medv,statistic = SE.fn,R=1000)
```

the average bootstrap estimate of SE is 
```{r echo=FALSE}
round(mean(SE_bootstrap$t),4)
```
which differs from the original estimate by 

```{r echo=FALSE}
round((100 * (mean(SE_bootstrap$t) - SE)/SE),2)
```
percent

```{r}
plot(SE_bootstrap)
```

### d.
Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of medv. compare it to the results obtained using t.test(Boston$medv).

```{r}
n <- nrow(Boston)
mu <- mean(Boston$medv)
SE <- mean(SE_bootstrap$t)

alpha <- 1 - 0.95
z <- qnorm(1 - alpha/2)
error <- z*SE

ttest <- t.test(Boston$medv)$conf.int
```

```{r echo=FALSE}
paste("using z stat, the 95% confidence interval is between"
      ,round(mu - error,4)
      ,"and"
      ,round(mu + error,4)
      )

paste("using 2*SE, the 95% confidence interval is between"
      ,round(mu - 2*SE,4)
      ,"and"
      ,round(mu + 2*SE,4)
      )

paste("using t.test, the 95% confidence interval is between"
      ,round(ttest[[1]],4)
      ,"and"
      ,round(ttest[[2]],4)
)
```
